---
title: "Spatial Assignment 3"
author: "M.C. Abbott, Manoel Pereira Neto, Jack Halverson"
date: "9/22/2021"
output: html_document
---

```{r, results = FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```

# Load and transform data
```{r}
cultural <- st_read("https://data.sfgov.org/api/geospatial/5xmc-5bjj?method=export&format=KML")
planning <- st_read("https://data.sfgov.org/api/geospatial/ttns-6zj3?method=export&format=KML")
buyouts <- st_read("https://data.sfgov.org/api/geospatial/wmam-7g8d?method=export&format=KML")
inclusionary <- st_read("Residential_Projects_With_Inclusionary_Requirements.kml")

CA_zone_3 <- "+proj=lcc +lat_1=38.43333333333333 +lat_2=37.06666666666667 +lat_0=36.5 +lon_0=-120.5 +x_0=2000000 +y_0=500000 +ellps=GRS80 +units=m +no_defs "

cultural <- cultural %>%
  st_transform(CA_zone_3)

planning <- planning %>%
  st_transform(CA_zone_3)

buyouts <- buyouts %>%
  st_transform(CA_zone_3)

inclusionary <- inclusionary %>%
  st_transform(CA_zone_3)
```

# Number and proportion of buyouts near inclusionary
```{r}
inclusionary_buffer <- st_buffer(inclusionary, dist = 400) %>%
  st_union()

ggplot(inclusionary_buffer) +
  geom_sf() +
  theme_map()


buyouts_inclusionary<- buyouts[inclusionary_buffer,]
  
ggplot(inclusionary_buffer) +
  geom_sf() +
  geom_sf(data = buyouts_inclusionary, 
          color = "darkgreen", 
          size = 0.01) +
  theme_map()


buyouts <- buyouts %>%
  st_join(buyouts_inclusionary) %>%
  mutate(by_inclusionary = !is.na(Description.y))


n_inclusionary_buyouts <- sum(buyouts$by_inclusionary)

n_inclusionary_buyouts


n_buyouts <- length(buyouts$by_inclusionary)

pct_inclusionary_buyouts <- n_inclusionary_buyouts / n_buyouts

pct_inclusionary_buyouts


left_side  <- st_bbox(planning)$xmin
top_side <- st_bbox(planning)$ymax

ggplot() +
  geom_sf(data = planning, fill = "light gray", color = "black") +
  geom_sf(data = buyouts, size = 1,
          aes(color = by_inclusionary)) +
  scale_color_manual(values = c("red", "blue"),
          name = "San Francisco 1% buyouts Projects \nby distance to a StreetSmbuyoutss Mural", 
          labels = c("No StreetSmbuyoutss Mural within 400 m",
                     "StreetSmbuyoutss Mural within 400 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side,
           label = paste("Of the ", 
                         prettyNum(n_buyouts, big.mark = ","),
                         " 1% buyouts Projects in San Francisco\n", 
                         prettyNum(n_inclusionary_buyouts, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_inclusionary_buyouts, digits = 0),
                         "%) are within 400\nmeters of a StreetSmbuyoutss Mural.",
                         sep = ""),
           hjust = 0, vjust = 1, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```


# Number and proportion of inclusionary near 1% buyouts projects
```{r, }
buyouts_buffer <- st_buffer(buyouts, dist = 400) %>%
  st_union()

ggplot(buyouts_buffer) +
  geom_sf() +
  theme_map()


inclusionary_buyouts <- inclusionary[buyouts_buffer,]
  
ggplot(buyouts_buffer) +
  geom_sf() +
  geom_sf(data = inclusionary_buyouts, 
          color = "darkgreen", 
          size = 0.01) +
  theme_map()


inclusionary <- inclusionary %>%
  st_join(inclusionary_buyouts) %>%
  mutate(by_buyouts = !is.na(Name.y))

n_buyouts_inclusionary <- sum(inclusionary$by_buyouts)

n_buyouts_inclusionary


n_inclusionary <- length(inclusionary$by_buyouts)

pct_buyouts_inclusionary <- n_buyouts_inclusionary / n_inclusionary

pct_buyouts_inclusionary


left_side  <- st_bbox(planning)$xmin
top_side <- st_bbox(planning)$ymax

ggplot() +
  geom_sf(data = planning, fill = "light gray", color = "black") +
  geom_sf(data = inclusionary, size = 1,
          aes(color = by_buyouts)) +
  scale_color_manual(values = c("red", "blue"),
          name = "Boston inclusionary\nby distance to a parking meter", 
          labels = c("No 1% buyouts within 30 m",
                     "1% buyouts within 30 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_inclusionary, big.mark = ","),
                         " trees in Boston\n", 
                         prettyNum(n_buyouts_inclusionary, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_buyouts_inclusionary, digits = 0),
                         "%) are within 30\nmeters of 1% buyouts.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```


# Number of buyouts in Planning Districts
```{r}
planning <- planning %>%
  mutate(num_buyouts = lengths(st_covers(planning, buyouts)))

ggplot(planning) +
  geom_sf(color = NA, 
          aes(fill = num_buyouts)) +
  scale_fill_viridis_c(name = "San Francisco planning districts\nby number of 1% buyouts Projects",
                       breaks = breaks <- seq(0, 50, by = 10),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "buyouts projects")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"), 
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```


# JACK - Average Number of Projects with Inclusionary Requirements in Planning Districts
```{r}
planning <- planning %>%
  mutate(num_inclusionary = lengths(st_covers(planning, inclusionary)))


right_side <- st_bbox(planning)$xmax
left_side  <- st_bbox(planning)$xmin
top_side <- st_bbox(planning)$ymax
bottom_side <- st_bbox(planning)$ymin

ggplot(planning) +
  geom_sf(color = NA, 
          aes(fill = num_inclusionary)) +
  coord_sf(xlim = c(left_side-1000, right_side+1000), 
           ylim = c(bottom_side-1000, top_side+1000), expand = FALSE) +
  scale_fill_viridis_c(name = "San Francisco Planning Districts\nby Projects with Inclusionary Requirements",
                       breaks = breaks <- seq(0, 120, by = 20),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "projects")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) +
  theme(legend.position = "right")
```

# Density of inclusionary in Planning Districts
```{r}
planning <- planning %>%
  mutate(area = set_units(st_area(planning), km^2)) %>%
  mutate(inclusionary_dens = as.numeric(num_inclusionary / area))

ggplot(planning) +
  geom_sf(color = NA, 
          aes(fill = inclusionary_dens)) +
    scale_fill_viridis_c(name =
                           "San Francisco Planning Districts\nby inclusionary density",
                       breaks = breaks <- seq(0, 100, by = 10),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "inclusionary per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

# JACK - Average distance of Projects with Inclusionary Requirements from Rental Buyouts
```{r}
inclusionary <- inclusionary %>%
  mutate(buyouts_dist = st_nn(inclusionary, buyouts, 
                           returnDist = TRUE)$dist) %>%
  mutate(buyouts_dist = as.numeric(buyouts_dist))


avg_buyouts_dist <- mean(inclusionary$buyouts_dist)

avg_buyouts_dist


right_side <- st_bbox(planning)$xmax
left_side  <- st_bbox(planning)$xmin
top_side <- st_bbox(planning)$ymax
bottom_side <- st_bbox(planning)$ymin

ggplot() +
  geom_sf(data = planning,fill = "light gray", color = "black") +
  geom_sf(data = inclusionary, size = 0.75,
          aes(color = buyouts_dist)) +
  coord_sf(xlim = c(left_side-1000, right_side+1000), 
           ylim = c(bottom_side-1000, top_side+3000), expand = FALSE) +
  scale_color_viridis_c(name = 
                          "San Francisco Projects with Inclusionary Requirements\nby distance to Rental Buyouts") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side - 500, 
           y = top_side + 500, 
           label = paste("On average, a San Francisco Project\nwith Inclusionary Requirements\nis ", 
                         prettyNum(avg_buyouts_dist, digits = 3),
                         " meters from a Rental Buyout.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) +
  theme(legend.position = "right")
```

# Number of inclusionary in Cultural Districts
```{r}
cultural <- cultural %>%
  mutate(num_inclusionary = lengths(st_covers(cultural, inclusionary)))

ggplot() +
  geom_sf(data = planning, fill = "light gray", color = "black") +
  geom_sf(data = cultural, color = NA, 
          aes(fill = num_inclusionary)) +
  scale_fill_viridis_c(name = "San Francisco planning districts\nby number of 1% buyouts Projects",
                       breaks = breaks <- seq(0, 10, by = 2),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "inclusionary")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```


# Number of buyouts in Cultural Districts
```{r}
cultural <- cultural %>%
  mutate(num_buyouts = lengths(st_covers(cultural, buyouts)))

ggplot() +
  geom_sf(data = planning, fill = "light gray", color = "black") +
  geom_sf(data = cultural, color = NA, 
          aes(fill = num_buyouts)) +
  scale_fill_viridis_c(name = "San Francisco planning districts\nby number of 1% buyouts Projects",
                       breaks = breaks <- seq(0, 15, by = 10),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "buyouts")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```




# Number and proportion of Planning Districts that have Cultural Districts
```{r}
planning <- planning %>%
  mutate(num_cultural = lengths(st_overlaps(planning, cultural))) %>%
  mutate(has_cultural = num_cultural > 0)

n_cultural_planning <- sum(planning$has_cultural)

n_cultural_planning


left_side  <- st_bbox(cultural)$xmin
top_side <- st_bbox(cultural)$ymax

ggplot(cultural) +
  geom_sf(fill = "lightblue", color = NA) +
  geom_sf(data = planning,
          aes(fill = has_cultural)) +
  scale_fill_manual(values = c("cornsilk1", "darkseagreen1"),
          name = "Boston Neighborhoods\nby presence of Cultural Districts", 
          labels = c("Neighborhood without\na Cultural District",
                     "Neighborhood with an\n Cultural District")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side - 1000, 
           label = paste(n_cultural_planning ,
                         "of San Francisco's", 
                         length(planning$Name),
                         "Planning Districts contain\nor overlap with", 
                         "Cultural Districts."),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```







# Density of 1% buyouts in Planning Districts
```{r}
planning <- planning %>%
  mutate(area = set_units(st_area(planning), km^2)) %>%
  mutate(buyouts_dens = as.numeric(num_buyouts / area))

ggplot(planning) +
  geom_sf(color = NA, 
          aes(fill = buyouts_dens)) +
    scale_fill_viridis_c(name =
                           "San Francisco Planning Districts\nby 1% buyouts density",
                       breaks = breaks <- seq(0, 1000, by = 100),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "1% buyouts Projects per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

